### 2022-05-30 paper/paper_draft.Rmd
updated introduction and added an abstract

---
title: "MASU Paper"
author: "William Hemstrom"
date: "11/21/2021"
output:
  word_document:
    reference_docx: "style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "Citations.bib"
csl: "council-of-science-editors-author-date.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, include = FALSE)
library(ggplot2); library(RefManageR); library(snpR)

```

#Abstract
Within Masu salmon (Oncorhynchus masou), individuals can exhibit one of two life cycles: resident or anadromous. The expresson of either life history is mainly influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Here, we explore if there are genetic factors that influence the development of either residency or anadromy. In order to look for patterns of relatedness among resident and anadromous individuals, we made a PCA chart, did association tests, and made a random forest (RF). Ultimately, we did not find any significant associated genetic elements, but further research with a larger sample size is needed in order to fully assess potential genetic influences on anadromy and residency.

#Introduction

<!-- Declines in migratory fishes -->
<!-- 	Trends and potential causes. -->

<!-- Importance of Masu -->
<!-- 	Vulnerability -->
<!-- 		Very prone to fishing pressure -->
<!-- 		Very limited distribution -->

<!-- Life History -->
<!-- 	What decides if an indivividual is a resident/anadromous? -->

<!-- Genetics of resident vs anadromous -->

<!-- What did we do and find? -->
<!-- 	Thesis -->

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in the recent years (CITE). In North America, for example,there have been significant decreases indiversity within pacific salmonid species due hydrologic and dam alterations (CITE), and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams (CITE). Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to these dams blocking the fishes' migratory paths (CITE- need a research paper for this number)

Masu salmon (Oncorhynchus masou, otherwise known as cherry salmon) have the most limited distribution among the Pacific salmon (CITE). Since they spend a greater portion of their life in freshwater rivers or streams than most other migratory salmon, they are particularly more exposed to fishing and sportfishing pressures (CITE).Their increased vulnerability makes it critical to understand the details that influence and control migratory life history variance in the species. Masu in particular (like Oncorhynchus mykiss, steelhead salmon/rainbow trout) are partially migratory, with large resident and anadromous populations (CITE). As is typical, anadromous masu are initially born in the rivers then migrate to the seas where they develop larger body sizes, then eventually migrate back to their natal rivers in order to breed (CITE). Resident masu in contrast sexually mature at a younger age and stay within the river systems throughout their entire lives (Morita and Nagasawa)

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Consequentially, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous (CITE). For example, X et al (YEAR) has shown that the number of resident males have increased over the last X years while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased, which has led to a significant decrease in the number of anadromous individuals overall.

In contrast, the genetic factors that contribute to migration in masu salmon have not been well studied.  Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas  crosses between anadromous individuals resulted in the most anadromous smolts. It is not surprising, therefore, that a recent study by Hess et al. (2016), on 15,239 single nucleotide polymorphisms (SNPs) found of three SNPs from a gene that was strongly linked with adult migration timing, and that 18 total SNPs that were connected to 60% of the trait variation. Despite this, however, growth rates are known to strongly influence anadromy in the species—individuals which grow faster as juveniles are far more likely to stay as residents than otherwise (CITE). Indeed, migration in the species seems to be controlled jointly the interplay of genetic and environmental factors:

The degree of genetic control of migration in masu, however, is still not well known. To improve our understanding of the mechanisms behind residency and anadromy in the species,  we sequenced a set of anadromous and resident masu individuals using high-resolution Restriction Associated Digest (RAD) sequencing.  However, we did not find any clear indicator of the genetic difference between ouranadromous and resident Masu salmon. While this may suggest that there is not that there is a strong genetic basis for Masu salmon migratory life history variation, further studies using a bigger sample size from a larger geographic range would need to be conducted in order to fully investigate the influences of genetics on Masu salmon migratory life cycles.

<!-- Anadromous salmon -> dams causing them to not be able to jump over dams -> is we explore the genetic influences of anadromy and residents in MASU salmon then we can tell where to focus our conservation efforts. -->

<!-- Parr size was generally larger and maturation was more common at more southern latitudes -->

<!-- Population density and temperature two most important factors affecting the growth of the fish.  -->

<!-- superior development promotes residency (early maturation) -->

<!-- need to do citations -->
```{r, include=FALSE, eval=TRUE}
dat<-read.table("../data/genotypes.geno")                        
meta<- read.table("../data/sample_meta.txt", header=TRUE)
snp.meta<- dat[,1:2]
colnames(snp.meta)<- c("chr", "position")
library(snpR)
# dat<- import.snpR.data(dat[,-c(1:2)], snp.meta = snp.meta, sample.meta = meta)
# dat<- filter_snps(dat, maf = 0.05, hwe = 0.000001)
# number_snps_after_filtering <- nsnps(dat)
# 
# #calculating genetic diversity
# dat <- calc_ho(dat, c("pop", "phen", "pop.phen"))
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="ho") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_fis(dat, c("pop", "phen"))
# dat <- calc_pi(dat, c("pop", "phen"))
# get.snpR.stats(dat, c("pop"), stats="pi") $weighted.means
# dat <- calc_private(dat, c("phen", "pop","pop.phen"))
# get.snpR.stats(dat, c("pop"), stats="private") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_pairwise_fst(dat, c("phen", "pop","pop.phen"), boot = 100, boot_par = 2)
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="fst") $weighted.means
# get.snpR.stats(dat, c("pop", "phen"), stats= c("ho", "pi", "fst")) $weighted.means
# 
# gp<- run_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)
# library(data.table)
# cv<- cross_validate_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)

# saveRDS(dat, "processed_genotypes.RDS")
# saveRDS(gp, "gp.RDS")

dat <- readRDS("../processed_genotypes.RDS")
gp <- readRDS("gp.RDS")

```
# Methods:
## Genotyping 
We extracted DNA from our samples using the magnetic bead approach [@Ali2016]. We then prepared Restriction Associated Digest (RAD) libraries using the Pst1 restriction enzyme [@Ali2016] and sequenced the results on an Illumina Hi-Seq 4000 using 150bp paired-end sequencing. We aligned the data to the O. mykiss v6 reference genome [@Pearse2019] using Burrows-Wheeler Alignment algorithm Aligner [@Li2009]. We then filtered the data for PCR duplicates and improper pairs using SAMtools [@Li2009]. Next, we called the genotypes by using the SAMtools genotype likelihood models [@Li2009] used in the ANGSD software package with a minimum mapping and base call quality score of 20, a SNP p-value of 1e-8, and a minimum acceptable minor allele frequency of 0.05 [@Korneliussen2014]. We also used a minor allele frequency filter (MAF) and a Hardy-Weinberg equilibrium filter using snpR when calling the genotypes [@Wigginton2005]. 
<!-- flagstat google (data script i wrote) -->

```{r, eval=TRUE, include=FALSE}
mat <- read.table("../angsdput.covMat")
IBS <- data.table::fread("../angsdput.covMat")
nrow(IBS)
IBS <- IBS[-32,-c(97, 32)]
pca_r <- eigen(IBS)
pca <- pca_r$vectors
write.table(pca[,1:20], "pca20.txt", col.names = FALSE, row.names = FALSE, quote=FALSE, sep="\t")
colnames(pca) <- paste0("PC", 1:ncol(pca))
meta <- read.table("../data/sample_meta.txt", header = TRUE)[-32,]
pca <- cbind(pca, meta)

loadings <- pca_r$values/sum(pca_r$values)
loadings <- round(loadings*100, 2)
pca$pop <- as.factor(pca$pop)
pca$phen <- as.factor(pca$phen)

library(ggplot2)
p1 <- ggplot(pca, aes(x = PC1, y = PC2, color = pop, shape = phen)) + theme_bw() +
  scale_color_viridis_d() + geom_point(size = 4) + 
  ggplot2::xlab(paste0("PC1 (", loadings[1], "%)")) + 
  ggplot2::ylab(paste0("PC2 (", loadings[2], "%)"))
p1
pca.centroids <- aggregate(pca[,1:2], list(stream = pca$pop), mean)
combine <- rbind(pca[, 1:2], pca.centroids[,-1])
dcombine<- dist(combine, upper = TRUE) 
dcombine<- as.matrix(dcombine)
dcombine<-dcombine[(nrow(dcombine) - nrow(pca.centroids) + 1):nrow(dcombine), -((nrow(dcombine) - nrow(pca.centroids) + 1):nrow(dcombine))]
dcombine <- t(dcombine) 
dcombine <- as.data.frame(dcombine)
dcombine$pop <- pca$pop
dcombine$phen <- pca$phen
colnames(dcombine)[1:2] <- c("CH", "TK")
dcombine$phen <- factor(dcombine$phen)
dcombine$pop <- factor(dcombine$pop)
dcombine$d <- 0
dcombine$d[dcombine$pop=="CH"] <- dcombine$TK[dcombine$pop=="CH"]
dcombine$d[dcombine$pop=="TK"] <- dcombine$CH[dcombine$pop=="TK"]
mod2 <- lm(d~phen, dcombine[-32,])
ggplot(dcombine, aes(x=pop, y=d, color=phen)) + geom_boxplot() + theme_bw() + scale_color_viridis_d(end = 0.5)

```

##  Diversity Stats
We calculated the observed heterozygosity (Ho), heterozygote/homozygote ratio per individual (Het/Hom), the average number of pairwise differences (pi), and fixation index (Fst) using the snpR R package [@Hohenlohe2010; @Wier1984; @Hemstrom2021].

```{r, eval=TRUE, include=FALSE}
# dat <- run_random_forest(dat, response = "phen", mtry = floor(nrow(dat)/2), num.trees = 100000)
# RF <- dat$models
# dat <- dat$data
# summary(RF$.base_.base)
# saveRDS("RF", "RF.RDS")
RF <- readRDS("RF.RDS")

association <- data.table::fread("../association_out_cov.lrt0.gz")
association$p <- pchisq(association$LRT, 1, lower.tail = FALSE)
plot_manhattan(association, "p", chr = "Chromosome", bp= "Position", log.p = TRUE)

```

## Association Testing
We performed a basic association test on the data using gmmat to correct for the related among individuals and population structure (). We then performed a random forest analysis in R-studio using the ranger R package [@Wright2017].  We used 100,000 trees and used mtry of half the number of snp's. We used a random forest to predict whether the samples would be resident or migratory. Using the BGLR program, we did a genomic prediction using BayesB distribution to estimate the heritability [@Meuwissen2001]. We used 1000 burn- ins and iterations=10,000. The heritability test was done to determine the degree in which phenotypic variation is due to genetic differences among the salmon populations.

## Population Structure and Migratory Relatedness
In order to detect any underlying population structuring within our stream systems, we used the software package NGSadmix to group samples into *k* clusters for each value of *k* between one and seven [@Skotte693]. To do so, we first called genotype posterior probailites using ANGSD with the same parameters above in addition to (fill in here). To assess variance in clustering within a given *k* value, we ran NGSadmix ten times for each value of *k* and combined the resulting data using CLUMPP [@Jakobsson2007]. We used the pophelper and snpR packages to handle files and produce plots [@Hemstrom2021; @Francis2017].

For comparison, we also conducted a Principle Component Analysis chart (PCA) using the ANGSD-IBS method [@Korneliussen2014]. To look for patterns in relatedness between resident and andromous individuals, we calculated the mean PC1 and PC2 coordinates for each stream, then calculated the distance between each individual sample the mean for the other population.

<!-- include that we did a one-way anova to determine if either phenotype was more closely related to the mean of the opposite population -->

```{r ngsadmix, include=FALSE}
setwd("../ngsadmix/")
ngsadmix <- plot_structure(".qopt", paste0(meta$pop, "_", meta$phen), facet.order = c("CH_A", "CH_R", "TK_A", "TK_R"), k = 1:7)
ngsadmix <- ngsadmix$plot + xlab("Population/Phenotype")
```

# Results 
```{r, include=FALSE}
MASU_data_table <- readr::read_delim("~/MASU_data_table", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE, skip = 4)
MASU_data_table <- as.data.frame(MASU_data_table)
MASU_data_table[,3] <- gsub("%", "", MASU_data_table[,3])
MASU_data_table[,5] <- gsub("%", "", MASU_data_table[,5])
MASU_data_table[,3]<- as.numeric(MASU_data_table[,3])
MASU_data_table[,5]<- as.numeric(MASU_data_table[,5])
mean(MASU_data_table[,3], na.rm = TRUE)
mean(MASU_data_table[,5], na.rm = TRUE)
```

##Genotyping 
The number of unfiltered reads are `r sort_read_count`.  After filtering, the number of reads were `r sort_flt_read_count`. We called a total of `r number_snps_after_filtering` snps and the number of IBS loci were `r nrow(IBS)`. 
```{r read, include=FALSE}
library(readr)
sort_read_count <- read.csv("~/Downloads/sort.read_count.txt", header = FALSE)
View(sort_read_count)
sort_read_count <- sum(sort_read_count$V1)

sort_flt_read_count <- read.csv("~/Downloads/sort.flt.read_count.txt", header = FALSE)
View(sort_flt_read_count)
sort_flt_read_count <- sum(sort_flt_read_count$V1)

```

flt read = `r sort_flt_read_count` 
<!-- flt read = 90449126 -->
unfiltered reads = `r sort_read_count` 
<!-- unfiltered reads = 163200129 -->
number of snps = `r number_snps_after_filtering`
Number of snps= 71175?
number of IBS loci = `r nrow(IBS)`

##  Diversity Stats
The mean Ho values are relatively similar between the anadromous and residents individuals and between the CH and TK populations. The fst values between the two populations and between anadromous and resident individuals are low, which indicated that there is not much genetic diversity between the two populations and two phenotypes. There are also no private alleles between the two populations, which also indicated low genetic diversity between the two populations of salmon.

## Association 
The random forest predicted 25 out of 52 anadromous individuals correctly and 17 out of 44 resident individuals correctly for a total prediction accuracy of 38.6%.

## Influence of Anadromy on Population Structure
We fit a simple linear regression with the distance to the other population cluster as the response variable and the migratory phenotype as the explanatory variable. We found that the anadromous individuals clustered signficiantly closer to the cluster center for individuals from the other population (one-way ANOVA, p-value = 0.0419). This indicates that the anadromous individuals within one population are more closely related to the individuals in the other population. 
<!-- The p value (phenR) is 0.03981775. The p value for the phenotypes is 0.0419.  -->

# Discussion 
## Results Discussion
According to the PCA, there is overlap between the anadromous and resident individuals from both populations. Although having overlapping clusters does not exclude the possibility of a genetic influence, our findings did not clearly show associated genetic elements between resident and anadromous fish. As mentioned in the previously to look for patterns in relatedness between resident and andromous individuals, we calculated the mean PC1 and PC2 coordinates for each stream, then calculated the distance between each individual sample and the mean for the other population. It is important to note that we did find that anadromous individuals were more closely related to individuals in the other population. 
<!-- This can potentially be attributed to where the samples were collected. For example, if the anadromous individuals from one stream were collected in an area that was geographically near the individuals in the other stream, this could explain this observation. (not sure how to interpret the metadata to confirm this) -->
We also made a manhattan plot and did not find many individuals that had a locus with a high p-value. This indicates that there is not an area in the genome that is strongly associated with the given trait (anadromy vs. resident). The random forest also did not have strong predictive power for anadromy and residency because it could not accurately group the individuals to their actual life cycle. 
Overall there is not strong evidence for a genetic influence that determines the MASU salmon's life cycle.

## Ramifications for MASU salmon
In these populations it is possible that residency and anadromy within MASU salmon are not solely genetica lly controlled and that we must take into consideration the environmental influences. When comparing MASU salmon to steelhead, steelhead and their life cycles are more heavily influenced by genetic factors. As of now, MASU salmon and their life cycles are mainly influenced my environmental factors, but further studies with larger sample sizes would need to be conducted in order to see if any genetic influences play a role. In other words, we would need still need to figure out if MASU salmon life cycles are influenced by the additive effects of genetics and environment, the multiplying effetcs of genetics and environment, or simply a result of purely environmental effects.

<!-- talk about genetic mechanism in steelhead and how that varies in environmental  -->
## Ramifications for Salmonids in General 
In terms of the importance of this work, figuring out if MASU salmon life cycles have genetics influences will go towards their conservation efforts. For example, within steelhead fish, anadromous individuals will have to migrate upstream in order to move towards the ocean. On some occurrences the anadromous steelhead will be stopped midway towards their journey to the ocean due to manmade dams. This has caused a reduction in diversity within these fish by causing an apparent increase in resident individuals within the population. Knowing whether a fish is anadromous or resident would be helpful because people can advocate for dams to built in areas where there are a lot of anadromous fish. This same logic can also be applied towards the conservation of MASU salmon.

<!-- control shift c --> 
[@Gowen2014; @Ali2016]
Gowen et al.[-@Gowen2014]